Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-01-29 Manuscript/manuscript.Rmd
manuscript edits

Genomic annotation data (GRCh37/hg19 human genome assembly) were obtained from the UCSC Genome Browser Database [@Tyner:2017aa] (Supplementary Table 1). The annotations consisted of a variety of different chromatin states, histone marks, and transcription factor binding sites using ChIP-seq peak experiments. Each genomic annotation was represented by a set of regions (their genomic coordinates) annotated as having a functional property or biological significance.

@@

The data sets used for modeling were composed of the full set of genomic bins that either overlapped with a flanked TAD boundary or did not. These two classes made up the minority and majority sets respectively. The data was then split into a 7:3 ratio of training to testing sets. Each training and testing set was composed of a similar majority to minority class ratio as the full data set. That is, for the task of prediction, the test set remained constant for all ensemble systems in order to account for a fair and unbiased comparison.

@@

3. **Log2 Distance**: For each genomic bin, the $log_{2}$ of the distance (in bases) from the center of the bin to the center of the nearest region of a respective genomic annotation region [??? Decide whether to use "annotation" or "feature"] was calculated. The log transformation was used as a normalization technique in order to account for the skewness of the data (Supplementary Figure 1).

To assess the predictive performance for each predictor type, we performed individual random forest classification algorithms utilizing each set predictor types as the feature space, in addition to a model with all combined types. Precision-Recall (PR) Curves, and subsequent area under the PR curves (AUPRC), were used to evaluate which predictor type contributed to the best model performance. Recursive feature elimination (RFE) was then used to determine the optimal number of features for models within each set of predictor types. In doing so, random forest models were used to assign weights to a set features of a given size. The optimal number of features was determined by the AUPRC of the corresponding model in order to reduce the feature space accordingly.

@@

### CLASSIFICATION

@@

### MODEL EVALUATION & PREDICTION

@@

Here, TP refers to the number of bins correctly identified as containing a TAD boundary (true positives), FP refers to the number of bins incorrectly identified as containing a TAD boundary (false positives), TN refers to the number of bins correctly identified as not containing a TAD boundary (true negatives), and FN refers to the number of bins incorrectly identified as not containing a TAD boundary. Each of these quantities are obtained from the confusion matrix created by validating the model on the testing set. The evaluation metrics listed above are all considered single threshold measures. That is, they are defined for individual thresholds that separated class prediction. Thererfore they are unable to give an overview of the potential range of performance as thresholds vary. A full list of the single threshold measures used (including additional metrics not reported), along with their respective formulas, are provided in Supplementary Table 2. In order to evaluate the trade-off between precision and recall, we also considered a threshold-free measure in form of the area under the precision recall curve (AUPRC).

>>>>

Genomic annotation data (GRCh37/hg19 human genome assembly) were obtained from the UCSC Genome Browser Database [@Tyner:2017aa] (Supplementary Table 1). The annotations consisted of a variety of different chromatin states, histone marks, and transcription factor binding sites using ChIP-seq peak experiments for the GM12878, K562, ... cell lines. Each genomic annotation was represented by a set of regions (their genomic coordinates) annotated as having a functional property or biological significance.

@@

The data sets used for modeling were composed of the full set of genomic bins that either overlapped with a flanked TAD boundary or did not. These two classes made up the minority and majority sets respectively. The data was then split into a 7:3 ratio of training to testing sets. Each training and testing set was composed of a similar majority to minority class ratio as the full data set. For the task of performance evaluation and prediction, the test set remained constant for all ensemble systems to account for a fair and unbiased comparison.

@@

3. **Log2 Distance**: For each genomic bin, the $log_{2}$ of the distance (refered to hereafter simply as $\textit{distance}$), in basepairs, from the center of the bin to the center of the nearest region of a respective genomic annotation region [??? Decide whether to use "annotation" or "feature"] was calculated. The log transformation was used as a normalization technique in order to account for the skewness of the data (Supplementary Figure 1).

@@

### CLASSIFICATION & CROSS VALIDATION

@@

### PERFORMANCE METRICS

@@

Here, TP refers to the number of bins correctly identified as containing a TAD boundary (true positives), FP refers to the number of bins incorrectly identified as containing a TAD boundary (false positives), TN refers to the number of bins correctly identified as not containing a TAD boundary (true negatives), and FN refers to the number of bins incorrectly identified as not containing a TAD boundary. Each of these quantities are obtained from the confusion matrix created by validating the model on the testing set. The evaluation metrics listed above are all considered single threshold measures. That is, they are defined for individual thresholds that separated class prediction. Thererfore they are unable to give an overview of the potential range of performance as thresholds vary. In order to evaluate the trade-off between precision and recall, we also considered a threshold-free measure in form of the area under the precision recall curve (AUPRC).

###MODEL ASSESSMENT AND PREDICTION

We first assessed predictor type performance using random forest classification algorithms by utilizing each set of predictor types as the feature space, in addition to a model with all combined types. Models were evaluated across each combination of resampling techniques applied to the training set, for each cell line. Performance was based on the AUPRC

To assess the performance for each predictor type, we performed individual random forest classification algorithms utilizing each set of predictor types as the feature space, in addition to a model with all combined types,. Precision-Recall (PR) Curves, and subsequent area under the PR curves (AUPRC), were used to evaluate which predictor type contributed to the best model performance. Recursive feature elimination (RFE) was then used to determine the optimal number of features for models within each set of predictor types. In doing so, random forest models were used to assign weights to a set features of a given size. The optimal number of features was determined by the AUPRC of the corresponding model in order to reduce the feature space accordingly.
